---
title: "Practice101"
author: "Hugo van den Belt"
date: "2024-10-28"
output: html_document
---

#R Practice
This is just some random practice


## RMarkdown formatting
```{r}
# Header 1
## Header 2
### Header 3

# Bold: **bold text** or __bold text__
# Italic: *italic text* or _italic text_
```

## R Markdown kb shortcuts

* Insert a new code chunk: Cmd + Option + I
* Run the current chunk: Cmd + Shift + Enter
* Run all chunks above: Cmd + Option + P
* Comment/Uncomment line or selection: Cmd + Shift + C
* Knit document: Cmd + Shift + K


## R Syntax Notes
c() = concatenate
  Basically the same as making a Tuple in python but in R they call it a vector;
    - Computationally efficient
    - They're ordered
    - can have duplicates
    However, vectors are
    - mutable
    - meant for elements of the same type
      - R will convert all to character type if a string is in the vector:
      
```{r}
mixed <- c(1, "hello", TRUE)  # becomes: "1" "hello" "TRUE"

print(mixed)
```

list() = really just a python list

%>% = pipe variable   
  Perform multiple operations on a variable. Pretty neat



##Practice
     
First import libraries

```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
library(lubridate)
```


create synthetic data for sales for example

```{r}
sales_data <- data.frame(
  date = seq(as.Date("2023-01-01"), as.Date("2023-12-31"), by = "day"),
  revenue = runif(365, 1000, 5000),  # Random revenues between 1000 and 5000
  category = sample(c("Electronics", "Clothing", "Food"), 365, replace = TRUE)
)
```


aggregation with dplyr grouping by month and category, then calculate various summary statistics
```{r}
monthly_summary <- sales_data %>%
  mutate(month = floor_date(date, "month")) %>%  # Convert dates to monthly buckets
  group_by(month, category) %>%
  summarise(
    total_revenue = sum(revenue),
    avg_daily_revenue = mean(revenue),
    n_days = n()
  )
```


this is how you make a function in R
```{r}
custom_function <- function(x, power = 2) {
  result <- x^power
  return(result)
}
```


different data structrures
```{r}
vector_practice <- c(1:100)  # numeric vector
matrix_practice <- matrix(1:25, nrow = 5)  # x5 matrix
list_practice <- list(
  numbers = 1:5,
  text = c("a", "b", "c"),
  mixed = list(1, "x", TRUE)  # Nested list with different data types
)
```

make plot with ggplot
```{r}
basic_plot <- ggplot(monthly_summary, aes(x = month, y = total_revenue, color = category)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Monthly Revenue by Category")

print(basic_plot)
```


correlation and eigenvalues
```{r}
correlation_matrix <- cor(matrix(rnorm(100), ncol = 4))  # Create correlation matrix
eigenvalues <- eigen(correlation_matrix)$values  # Calculate eigenvalues

print(correlation_matrix)
print(eigenvalues)
```


You can create random values without an additional library 
```{r}
simulated_data <- data.frame(
  x = rnorm(1000),  # Generate random x values
  y = rnorm(1000)   # Generate random y values
)

print(simulated_data)
```















