---
title: "Practice Week 1"
author: "Hugo van den Belt"
date: "2024-12-14"
output: html_document
---


## R Plotting Setup Notes

Basic visualization setup in R using two key libraries:
- `ggplot2`: Main graphics package for creating plots
- `kableExtra`: Package for enhanced table formatting in documents

The command `theme_set(theme_gray(15))` sets default plot styling with gray background and 15pt font. Other useful themes include `theme_minimal()`, `theme_bw()`, `theme_classic()`.

**Note: Include `{r plots_and_figures, include=FALSE}` at start of code chunk to run setup without showing code in final document.**
```{r plots_and_figures, include=FALSE}
library(ggplot2)
library(kableExtra)

theme_set(theme_gray(15))
```


## viewing DataFrames
reading a flat file
```{r}
CEO_Diary <- read.csv("/Users/hugovandenbelt/Documents/GitHub/GROUP_ASSIGNMENT/survey_response_data.csv")
```

head item (4 rows)
```{r}
head(CEO_Diary, 4)
```

can also tail
```{r}
tail(CEO_Diary, 10)
```


## Subsetting
You can subset a df very easily using square brackets, with the input `[rows, columns]`

```{r}
CEO_Diary[1:10,c(1:5,37, 39:42)] # Shows rows 1 through 10, and colums 1 through 5, 37, and 39 through 42
```

`:` slices (includes lower and upper limit)

The `apply()` function lets you apply a function across a data frame or matrix, either by rows or columns. It takes three main arguments:

1. The data frame/matrix (in this case, CEO_Diary)
2. A number indicating direction (1 for rows, 2 for columns)
3. The function to apply (in this case, class)

with our df this would look like:

```{r}
apply(CEO_Diary, 2, class)
```

and with a simpler df it could look like this:

```{r}
# sample data frame
sample_df <- data.frame(
    name = c("Alice", "Bob"),        
    age = c(25, 30),              
    active = c(TRUE, FALSE)        
)

# Apply class function to each column
apply(sample_df, 2, class)
```

For some reason it identifies all columns as characters. Why?

Apparently, when we apply the `apply()` function, R converts our df into a matrix, but in a matrix every element must be of the same type. We can use the `sapply()` function to see the classes directly:

```{r}
sapply(sample_df, class)
sapply(CEO_Diary, class)
```

## Some statistics

we can get the number of rows 

```{r}
nrow(CEO_Diary)
```

let me guess, we can get the number of columns by doing ncolumn or something

```{r}
ncol(CEO_Diary) # indeed
```


We can get summary statistics

```{r}
summary(CEO_Diary[1:10]) # Columns 1 through 10
```


Create a frequency table of a certain column
```{r}
table(CEO_Diary$finance)
table(CEO_Diary$n_functions)
```
The `$` operator selects columns


Building on this, we can view proportions
```{r}
prop.table(table(CEO_Diary$type))
```


we can visualise the proportions with a barplot

```{r}
barplot(prop.table(table(CEO_Diary$n_functions)))
```


## MODELING TIME

we can fit a linear model with chosen variables
```{r}
fit <- glm(strategy ~ consultants + politicians, data=CEO_Diary)
summary(fit)
```

The presence of consultants *increases* the probability of the meeting being about strategy. 
The presence of politicians *decreases* the probability of the meeting being about strategy. 

We can add as many independents as we'd like
```{r}
fit_all <- glm(strategy ~ consultants + politicians + journalists + unions + investors + lawyers, data=CEO_Diary)

summary(fit_all)
```

